---
title: "Random effects meta-analysis for correlated test statistics"
author:
- name: "[Gabriel Hoffman](http://gabrielhoffman.github.io)"
  affiliation: | 
    Icahn School of Medicine at Mount Sinai, New York

output:
  rmarkdown::html_document:
    highlight: pygments
    toc: false
    toc_depth: 3
    fig_width: 5
bibliography: ../inst/REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{remaCor: Random effects meta-analysis for correlated test statistics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, results="hide"}
knitr::opts_chunk$set(tidy=FALSE, cache=TRUE,
                      dev="png",
                      package.startup.message = FALSE,
                      message=FALSE, error=FALSE, warning=TRUE)
```	

<style>
body {
text-align: justify}
</style>

Standard approaches to meta-analysis assumes that effect sizes are statistically independent.  Here we provide methods for fixed and random effects meta-analysis when the correlation between effect sizes are known.

### Fixed effects meta-analysis

`LS()` implements fixed effect meta-analysis for correlated test statistics using method of @lin2009meta.  By default, correlation is set to identity matrix to for independent test statistics.  

### Random effects  meta-analysis

`RE2C()` implements random effect meta-analysis for correlated test statistics that jointly tests deviation of the mean from zero as well as effect size heterogenity.  This method uses the RE2 method of @han2011random, or RE2 for correlated test statistics from @han2016general.   By default, correlation is set to identity matrix to for independent test statistics.  (In addition, this function computes the two step RE2C method of @lee2017increasing to further test for heterogenity in effect size after applying a fixed effect test.)


  * stat1: statistic testing effect mean

  * stat2: statistic testing effect heterogeneity

  * RE2Cp: RE2 p-value accounting for correlelation between tests. (This is the p-value appropriate for most questions)

  * RE2Cp.twoStep: two step RE2C test after fixed effect test.
            Only evaluated if twoStep==TRUE.  (not typically used)

  * QE: test statistic for the test of (residual) heterogeneity

  * QEp: p-value for the test of (residual) heterogeneity

  * Isq: I^2 statistic

    QE, QEp and ISq are only evaluted if correlation is diagonal



### Examples
```{r rema}
library(remaCor)
library(metafor)

# Generate effects
library(mvtnorm)
library(clusterGeneration )

n = 4
Sigma = cov2cor(genPositiveDefMat(n)$Sigma)
beta = t(rmvnorm(1, rep(0, n), Sigma))
stders = rep(.1, n)

# Standard fixed effects meta-analysis
# of independent effects with metafor pacakge
rma( beta, sei=stders, method="FE")

# Standard random effects meta-analysis
# of independent effects with metafor pacakge
rma( beta, sei=stders, method="REML")

# Run fixed effects meta-analysis, assume identity correlation  
# Use Lin-Sullivan method
LS( beta, stders)

# Run fixed effects meta-analysis, accounting for correlation  
# Use Lin-Sullivan method
LS( beta, stders, Sigma)

# Run random effects meta-analysis, assume identity correlation  
RE2C( beta, stders)

# Run random effects meta-analysis, accounting for correlation 
RE2C( beta, stders, Sigma)
```

```{r out, echo=FALSE}
RE2C( beta, stders, Sigma)
```





# Session info
<details>
```{r sessionInfo}
sessionInfo()
```
</details>

## References
